### 2012-11-19 doc/semilagr/semilagrangian.tex

\documentclass[11pt]{article}

\usepackage{algorithm,algorithmic}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}
\usepackage{xspace}

\newcommand{\pforest}{\texttt{p4est}\xspace}

\setlength{\emergencystretch}{20pt}

\title{A scalable semi-Lagrangian method on octree meshes}
\author{}

\begin{document}

\maketitle

\begin{abstract}
We present a semi-Lagrangian method for solving the advection equation
on nested non-conforming hexahedral meshes.
\end{abstract}

\section{Introduction}

\section{The semi-Lagrangian method}

\begin{algorithm}
  \caption{Basic semi-Lagrangian procedure}
  \begin{algorithmic}
    \STATE Do something.
  \end{algorithmic}
\end{algorithm}

\end{document}