### 2015-06-04 paper/tex/abstract.tex
starting revisions ...

%!TEX root = draft.tex
\begin{abstract}
We present scalable parallel algorithms for the level-set method on adaptive Quadtree and Octree Cartesian grids. The algorithms are based on a domain decomposition technique and implemented using \texttt{MPI} and the open-source \texttt{p4est} library. An important contribution is a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions. This is achieved by introducing a scalable global interpolation scheme on adaptive tree-based grids. Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation. Both parallel algorithms scale on the Stampede supercomputer, where we are limited to 4096 cores currently. Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detail that would be impossible to achieve without a parallel adaptive strategy. We believe that the algorithms presented in this article will be of interest and useful to researchers working with the level-set framework and modeling multi-scale physics in general.
\end{abstract}
\begin{keyword}
Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep Semi-Lagrangian Method \sep Level-set Method
\end{keyword}
\maketitle