### 2016-06-20 doc/paper/tex/draft.tex

\documentclass[final]{elsarticle}
%\usepackage[pdftex,breaklinks,linktocpage,pagebackref,hyperindex,hyperfigures]
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{pdfsync}
% \usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{color}
\usepackage{caption}
\usepackage{url}
\usepackage{tikz}
\usepackage{diagbox}
\usepackage{kpfonts}

\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}
\DeclareMathSymbol{\vardiamond}{\mathalpha}{extraup}{87}

\hypersetup{
    colorlinks=true,      		% false: boxed links; true: colored links
    linkcolor=blue,       		% color of internal links
    citecolor=blue,       		% color of links to bibliography
    filecolor=black,      		% color of file links
    urlcolor=red,       		% color of external links
    pdffitwindow=true,
    pdfpagelayout=SinglePage
}

% defining some new commands
\newcommand{\etal}{\textit{et al.\ }}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\mohammad}[1]{\textcolor{blue}{{Mohammad: #1}}}
\newcommand{\tensor}[1]{\underline{\underline{\boldsymbol{#1}}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ben}{\begin{equation*}}
\newcommand{\een}{\end{equation*}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\bean}{\begin{eqnarray*}}
\newcommand{\eean}{\end{eqnarray*}}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pdn}[3]{\frac{\partial^#1 #2}{\partial #3^#1}}
\newcommand{\noi}{\noindent}
\newcommand{\ind}{\indent}
\newcommand{\divg}[1]{\nabla \cdot \left(#1\right)}
\newcommand{\ddivg}[1]{\nabla \cdot #1}
\newcommand{\non}{\nonumber}
\newcommand{\fcite}[1]{(\textcolor{blue}{cite: #1})}
% \newcommand{\drawcircle}[2]{\protect  \tikz \protect \node[circle,draw,scale=#1,fill=#2] {};}
% \newcommand{\drawsquare}[2]{\protect  \tikz \protect \node[rectangle,draw,scale=#1,fill=#2] {};}
% \newcommand{\drawdiamond}[2]{\protect \tikz \protect \node[rectangle,draw,scale=#1,fill=#2,rotate=45] {};}
\newcommand{\drawcircle}[1]{\textcolor{#1}{$\bullet$}}
\newcommand{\drawsquare}[1]{\textcolor{#1}{$\blacksquare$}}
\newcommand{\drawdiamond}[1]{\textcolor{#1}{$\vardiamond$}}
\newtheorem{property}{Property}

\begin{document}

\title{Parallel Level-Set Methods on Adaptive Tree-Based Grids}

\cortext[cor]{Corresponding author: m.mirzadeh@engineering.ucsb.edu}
\address[MECHE]{Department of Mechanical Engineering, University of California, Santa Barbara, CA 93106-5070, United States.}
\address[CS]{Department of Computer Science, University of California, Santa Barbara, CA 93106-5110, United States.}
\address[UBONN]{Institute for Numerical Simulation, University of Bonn, 53115 Bonn, Germany.}

\author[MECHE]{Mohammad Mirzadeh\corref{cor}} \author[MECHE]{Arthur Guittet} \author[UBONN] {Carsten Burstedde} \author[MECHE,CS]{Frederic Gibou}

\input{abstract}
\input{introduction}
\input{levelset_method}
\input{parallel_algorithms}
\input{scaling_results}
\input{applications}
\input{conclusion}

\section*{References}
\bibliographystyle{plain}
\addcontentsline{toc}{section}{\refname}
\bibliography{casl,refs,bonn}

\end{document}